---
output: github_document
---

<!-- Example4.md is generated from Example4.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 4: time series with monthly varying variance

### 1. Simulate a time series with one different variance every calendar month 

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# time series simulation function
# Note: by convention the position of a change-point is the last point in the segment
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]
  time_series <- cumsum(changes)

  noise <- rnorm(n = length_series, mean = 0, sd = 1)
  sd = noise_stdev[as.numeric(format(mydate, "%m"))]
  time_series <- time_series + noise * sd

  return(time_series)
}

# specify the simulation parameters
n = 1000                            # length of time series
cp_ind <- c(200, 600)               # position of change points (index in time series)
segmt_mean <- c(-1, 1, 2)           # mean value of segments
noise_stdev <- c(0.1, 0.3, 0.7, 1.2, 1.8, 2, 2, 1.8, 1.2, 0.7, 0.3, 0.1) # 12 values, one per month (Jan to Dec)
set.seed(1)                         # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df = data.frame(date = mydate, signal = mysignal)
plot(df$date, df$signal, type = "l",xlab ="Date",ylab="signal")
abline(v = mydate[cp_ind], col = "red", lty = 2)
```

### 2. Segmentation

Run the segmentation with default parameters:

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
str(seg)
```

```{r}
seg$Tmu
```

### 3. Visualization of the time series with segmentation results superposed

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

Note: the segmentation works well despite the steep variations in the noise variance

### Compare estimated noise variance vs truth
```{r}
seg$MonthVar - noise_stdev ** 2
sd(seg$MonthVar - noise_stdev ^ 2)^2
```

Note: increasing the length of the time series (more years) would improve the accuracy of the noise variance estimation
