---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```


# PMLSeg

PMLseg is a R package for the segmentation of univariate time series based on Penalized Maximum Likelihood. The method detects changes in the mean signal in the presence of a periodic bias and variance changing over prescribed intervals. A preliminary version was published under the name `GNSSseg` which was developed for the detection of change-points in daily GNSS minus reanalysis IWV time series with annual periodic bias and variance changing on monthly intervals. PMLseg is more versatile and can be used with various modelling options (with or without periodic bias, constant or monthly varying noise) and algorithmic evolutions regarding the iterative inference scheme. The time series should be provided in a data frame containing two columns: date and signal. The package includes the segmentation function as well as auxiliary functions for the validation with metadata, cluster screening, statistical test of change-points, and visualization of the time series with segmentation and validation results. The theoretical basis of the method can be found in [@Quarello2022].
 
## Table of Contents

- [Installation](#installation)
- [Functions](#functions)
- [Examples](#examples)

## Installation

The development version of `PMLSeg` can be installed from GitHub with:

```{r, eval=FALSE}

# Install devtools, tidyr, and purrr if you haven't already
install.packages("devtools", dependencies = TRUE))
install.packages("tidyr", dependencies = TRUE))
install.packages("purrr", dependencies = TRUE))

# Download and install gfpop from
https://cran.r-project.org/src/contrib/Archive/gfpop/gfpop_1.1.1.tar.gz
install.packages("gfpop_1.1.1.tar.gz")

# Install PMLseg from GitHub
devtools::install_github("khanhninhnguyen/PMLSeg")
```

# Functions

The package contains the following functions:

```{r, eval=FALSE}
Segmentation: estimates the position of change-points in a time series

Validation: finds the closest metadata events for each change-point

Cluster_screening: detects clusters of change-points (based on a statistical test)

Test_CP: tests significance of every change-point

UpdateTimeSeries: removes data from the time-series in the clusters

UpdatedParametersForFixedCP: removes change-points in the clusters

PlotSeg: plots time series, change-points, metadata and validation results
```

Usage and help on functions can be found by ?"function" in R console, e.g.

```{r, eval=FALSE}
?Segmentation
```

## Example

Below we provide an example of the main functions with a simulated time series. More examples are given in the `Examples.md/` folder. 

### Simulation of a time series with changes in the mean and homogeneous variance

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)
library(purrr)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n <- 1000                    # length of time series
cp_ind <- c(200, 600)       # position of CPs (index in time series)
segmt_mean <- c(-1, 1, 2)   # mean value of segments
noise_stdev <- 1             # noise std dev (identical for all months)
set.seed(1)                 # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df <- data.frame(date = mydate, signal = mysignal)
```

### Segmentation

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE,
                   VarMonthly = FALSE)
str(seg)
```

The `Tmu` dataframe contains, for each segment: the index and date/time of beginning and end, the estimated mean `mean` and its standard erreor `se`, and the number of valid (non-NA) data points `np` in the signal:

```{r}
seg$Tmu
```

The time series with segmentation results superposed can be plotted with the `PlotSeg` function:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

The plot shows the signal (grey line), the estimated means (red line), the estimated noise std (cyan line). 
The y-intercept of the noise is the lower black line. Dashed vertical lines show the estimated times of the change-points.

### Validation of detected change-points with metadata

Metadata is represented by a data frame with 2 columns: `date`, `type`.

For the example, we create a fake metadata data frame with the true position of change-points:

```{r}
meta_ind = cp_ind               # index in time series of metadata information
meta_date <- df$date[meta_ind]  # corresponding date 
meta_type <- c("R", "RAD")      # type of information, e.g. R = receiver change, A = antenna change, D = radome change
metadata = data.frame(date = meta_date, type = meta_type)
metadata
```

Validate estimated change-point positions wrt metadata:

```{r}
valid_max_dist = 10             # maximum distance wrt metadata for a CP to be validated
valid = Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```
Note: valid$Distance gives the distance between estimated CP and metadata

Plot with metadata and validation results:

```{r}
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = FALSE, Metadata = metadata, Validated_CP_Meta = valid)
```

Metadata are indicated by grey symbols near the top of the figure.
Validated change-points are indicated by a red squares at the bottom of the figure.

### Test the significance of the detected change-points

Our segmentation method is based on maximum likelihood. Other methods use tests. Some users are more familiar with the concept of significance provided by such methods.
Since we have implemented a test in the cluster screening function (see Example 3), we found it useful to provide a test function which can be used in a post-segmentation step to evaluate the significance of the detected change-points.
In practice, it comes out that the change-points detected by PMLseg are almost always significant at the level 0.05.

```{r}

test <- Test_CP(Tmu = seg$Tmu, alpha = 0.05, detail = TRUE)
test
```

The test structure provides the list of updated change-points `UpdatedCP`, the variable `ChangeCP` indicates if the list has been changed compared to the input vector `Tmu`. An optinal variable `detail` is returned if the input option `detail` is TRUE. It provides the input information on the pairs of segments to the left and to the right of each change-point which have been used for the test, and the test outputs, `tstat`, `pval`, and `signif`, where `signif` = 1 means that a change-point is significant (this is the case when `pval < alpha`).

### Other examples 

See the other examples with more complex signals in the `Examples.md/` folder:

 - Example 1: time series with 2 change-points and constant noise
 - Example 2: time series with missing data
 - Example 3: time series with clusters of CPs
 - Example 4: time series with monthly varying variance
 - Example 5: time series with periodic bias
 - Example 6: time series with periodic bias, monthly variance, and gaps
 - Example 7: Explore the different penalty criteria

## References







