---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```


# PMLSeg

PMLseg is a R package for the segmentation of univariate time series based on Penalized Maximum Likelihood. The method detects changes in the mean signal in the presence of a periodic bias and variance changing over prescribed intervals. A preliminary version was published under the name `GNSSseg` which was developed for the detection of change-points in daily GNSS minus reanalysis IWV time series with annual periodic bias and variance changing on monthly intervals. PMLseg is more versatile and can be used with various modelling options (with or without periodic bias, constant or monthly varying noise) and algorithmic evolutions regarding the iterative inference scheme. The time series should be provided in a data frame containing two columns: date and signal. The package includes the segmentation function as well as auxiliary functions for the validation with metadata, cluster screening, statistical test of change-points, and visualization of the time series with segmentation and validation results. The theoretical basis of the method can be found in [@Quarello2022].
 
## Table of Contents

- [Installation](#installation)
- [Functions](#functions)
- [Examples](#examples)

## Installation

The development version of `PMLSeg` can be installed from GitHub with:

```{r, eval=FALSE}

# Install devtools, tidyr, and purr if you haven't already
install.packages("devtools")
install.packages("tidyr")
install.packages("purr")

# Download and install gfpop from
https://cran.r-project.org/src/contrib/Archive/gfpop/gfpop_1.1.1.tar.gz
install.packages("gfpop_1.1.1.tar.gz")

# Install PMLseg from GitHub
devtools::install_github("khanhninhnguyen/PMLSeg")
```

# Functions

The package contains the following functions:

```{r, eval=FALSE}
Segmentation(
  OneSeries,            # time series (data frame with 2 columns: $signal, $date in Date format)
  lmin = 1,             # minimum length of the segments
  Kmax = 30,            # maximal number of segments
  selectionK = "BM_BJ", # penalty criterion used for the model selection
  FunctPart = TRUE,     # if TRUE include a functional part (Fourier Series) to model the periodic bias
  selectionF = FALSE,   # if TRUE select the significant coefficients of the functional part
  initf = "ascendent",  # initialization mode of the iterative inference procedure
  VarMonthly = TRUE     # if TRUE use of monthly variance model, otherwise a constant variance.
)

Validation(
  OneSeries,            # time series data frame
  Tmu,                  # estimated change-point information, output from Segmentation function
  MaxDist = 62,         # maximum distance (in days) between detected CP in Tmu and known events in Metadata
  Metadata              # data frame with two columns: $date and the $type
)

Cluster_screening(
  Tmu,                  # estimated change-point information
  MaxDist = 80,         # maximal number of days between change-points within a cluster
  alpha = 0.05,         # significance level of test
  detail = FALSE        # if TRUE the output contains a $detail field with additional information on the test
)

Test_CP(
  Tmu,                  # estimated change-point information
  alpha = 0.05,         # significance level of test
  detail = FALSE        # if TRUE the output contains a $detail field with additional information on the test
)

UpdatedParametersForFixedCP(
  OneSeries,            # time series data frame
  ResScreeningTest,     # output from Cluster_screening function or Test_CP function
  FunctPart = TRUE,     # if TRUE update functional part
  selectionF = FALSE,   # if TRUE select significant coefficients in the functional
  VarMonthly = TRUE     # if TRUE use of monthly variance model, otherwise a constant variance
)

PlotSeg(
  OneSeries,            # time series data frame
  SegRes,               # output from Segmentation function
  FunctPart = TRUE,     # if TRUE plot functional
  Metadata = NULL,      # if not NULL plot metadata
  Validated_CP_Meta = NULL, # if not NULL plot validation result
  labelx = "date",
  labely = "signal",
  title = "Segmented time series"
)
```

Usage and help on functions can be found by ?"function" in R console, e.g.

```{r, eval=FALSE}
?Segmentation
```

## Examples

Several examples are given in the `Examples.md/` folder. Below we provide an excerpt of example1.md

### Simulation of a time series with changes in the mean and homogeneous variance

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)
library(purrr)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n <- 1000                    # length of time series
cp_ind <- c(200, 600)       # position of CPs (index in time series)
segmt_mean <- c(-1, 1, 2)   # mean value of segments
noise_stdev <- 1             # noise std dev (identical for all months)
set.seed(1)                 # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df <- data.frame(date = mydate, signal = mysignal)
```

### Segmentation

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
str(seg)
```

The `Tmu` dataframe contains, for each segment: the index and date/time of beginning and end, the estimated mean `mean` and its standard erreor `se`, and the number of valid (non-NA) data points `np` in the signal:

```{r}
seg$Tmu
```

The time series with segmentation results superposed can be plotted with the `PlotSeg` function:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

The plot shows the signal (grey line), the estimated means (red line), the estimated noise std (cyan line). 
The y-intercept of the noise is the lower black line. Dashed vertical lines show the estimated times of the change-points.

### Validation of detected change-points with metadata

Metadata is represented by a data frame with 2 columns: `date`, `type`.

For the example, we create a fake metadata data frame with the true position of change-points:

```{r}
meta_ind = cp_ind               # index in time series of metadata information
meta_date <- df$date[meta_ind]  # corresponding date 
meta_type <- c("R", "RAD")      # type of information, e.g. R = receiver change, A = antenna change, D = radome change
metadata = data.frame(date = meta_date, type = meta_type)
metadata
```

Validate estimated change-point positions wrt metadata:

```{r}
valid_max_dist = 10             # maximum distance wrt metadata for a CP to be validated
valid = Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```
Note: valid$Distance gives the distance between estimated CP and metadata

Plot with metadata and validation results:

```{r}
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = FALSE, Metadata = metadata, Validated_CP_Meta = valid)
```

Metadata are indicated by grey symbols near the top of the figure.
Validated change-points are indicated by a red squares at the bottom of the figure.

### Other examples 

See the other examples with more complex signals in the `Examples.md/` folder.

## References







