---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```


# PMLSeg

This package used to detects change-points in daily climatic time series modeled by piecewise mean, periodic bias, and monthly variance. Additionally, it includes functions to validate and visualize the segmentation results.

## Table of Contents

- [Installation](#installation)
- [Examples](#examples)

## Installation

You can install the development version of `PMLSeg` from GitHub with:

```{r, eval=FALSE}

# Install devtools if you haven't already
install.packages("devtools")

# Install the package from GitHub
devtools::install_github("khanhninhnguyen/PMLSeg")
```



## Examples

1. Creating a Time Series with Arbitrary Jumps

```{r}
library(PMLseg)

set.seed(1)
length_series = 1000
df = data.frame(date = seq.Date(from = as.Date("2010-01-01"),
                                to = as.Date("2010-01-01")+(length_series-1), 
                                by = "day"),  
                signal = rnorm(n = length_series, mean = 0, sd = 1))
plot(df1$date, df1$signal, type = "l")
head(df1, 3)
```

```{r}
# Adding 2 jumps 
# Function to generate jump series
generate_jump_series <- function(jump_indices, jump_amp, length_series) {
  jump_series <- rep(0, length_series)
  jump_indices <- c(1, jump_indices, length_series + 1)
  
  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- jump_amp
  
  jump_series <- cumsum(changes)
  
  return(jump_series)
}
# No cluster
jump_ind <- c(200, 600)
jump_amp <- c(0, 1, 1)
jump_series <- generate_jump_series(jump_ind, jump_amp, length_series)
df$signal <- df$signal + jump_series
plot(df$date, df$signal, type = "l")

```

2. Segmenting the Time Series

Adjust the `FunctPart` parameter based on whether the series includes seasonal mean components. This code run with default option for : `Kmax` = 30 (maximum number of segments), `selectionK` = "BM_BJ" (recommended penalized criterion), etc. 

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
str(seg)
```
Main results: The beginning, end, mean value, standard deviation, and number of points of segments are provided in the `Tmu` dataframe: 

```{r}
seg$Tmu
```

3. Visualizing Segmentation Results

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

4. Validating Detected Change-Points with Metadata

Change-points are validated if the segmentation finds them within  `MinDist` points from the metadata. Default `MinDist` is 62.

```{r}
# Create a metadata for example 
meta = data.frame(date = df$date[jump_ind],  
                type = c("example1", "example2"))
meta
```
```{r}
# Validate the segmentation result wrt metadata 
Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MinDist = 30 ,
           Metadata = meta)

```

5. Detecting Clusters of Change-Points

Check for clusters of close change-points, which may due to outliers. In this example, the last segment is shorter than the given threshold.

```{r}
# Validate the segmentation result wrt metadata 
screening = Cluster_screening(Tmu = seg$Tmu,
                              MaxDist = 80)

```

Update variables after removal of short segments:

```{r}
seg_upd = UpdatedParametersForFixedCP(OneSeries = df, 
                                      ResScreening = screening, 
                                      FunctPart = FALSE)
str(seg_upd)
```

Visualize the final results:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg_upd, 
        FunctPart = FALSE)
```








