---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```


# PMLSeg

PMLseg is a R package for the segmentation of univariate time series based on Penalized Maximum Likelihood. The method detects changes in the mean signal in the presence of a periodic bias and variance changing over prescribed intervals. It was initially developed for the detection of change-points in daily climate data with annual periodic bias and variance changing on monthly intervals. It is quite versatile and can be of more general-purpose usage. The package includes the segmentation function as well as auxiliary functions for the screening, validation and visualization of the segmentation results. The theoretical basis of the method was published in [@Quarello2022]. 
 
## Table of Contents

- [Installation](#installation)
- [Examples](#examples)

## Installation

The development version of `PMLSeg` can be installed from GitHub with:

```{r, eval=FALSE}

# Install devtools and tidyr if you haven't already
install.packages("devtools")
install.packages("tidyr")

# Download and install gfpop from
https://cran.r-project.org/src/contrib/Archive/gfpop/gfpop_1.1.1.tar.gz
install.packages("gfpop_1.1.1.tar.gz")

# Install PMLseg from GitHub
devtools::install_github("khanhninhnguyen/PMLSeg")
```



## Example 1

1. Simulate a time series with 2 change-points

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# time series simulation function
# Note: by convention the position of a change-point is the last point in the segment
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n = 1000                    # length of time series
cp_ind <- c(200, 600)       # position of change points (index in time series)
segmt_mean <- c(-1, 1, 2)   # mean value of segments
noise_stdev = 1             # noise std dev (identical for all months)
set.seed(1)                 # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df = data.frame(date = mydate, signal = mysignal)
plot(df$date, df$signal, type = "l",xlab ="Date",ylab="signal")
head(df, 3)
```

2. Segmentation

Run the segmentation with default parameters and no functional:

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
str(seg)
```

The `Tmu` dataframe contains, for each segment: the index of beginning and end, the esitmated mean and its standard erreor `se`, and the number of valid (non-NA) data points `np` in the signal:

```{r}
seg$Tmu
```

3. Visualization of the time series with segmentation results superposed

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

4. Validation of detected change-points with metadata

Metadata is represented by a data frame with 2 columns: `date`, `type`.

For the example, we create a fake metadata data frame with the true position of change-points:

```{r}
meta_ind = cp_ind               # index in time series of metadata information
meta_date <- df$date[meta_ind]  # corresponding date 
meta_type <- c("R", "RAD")      # type of information, e.g. R = receiver change, A = antenna change, D = radome change
metadata = data.frame(date = meta_date, type = meta_type)
metadata
```

Plot with metadata:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE, 
        Metadata = metadata) 
```

Validate estimated change-point positions wrt metadata:

```{r}
valid_max_dist = 62             # maximum distance wrt metadata for a CP to be validated
valid = Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```

Plot with metadata and validation results:

```{r}
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = FALSE, Metadata = metadata, Validated_CP_Meta = valid)
```

## References







