---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```


# PMLSeg

PMLseg is a R package for the segmentation of univariate time series based on Penalized Maximum Likelihood. The method detects changes in the mean signal in the presence of a periodic bias and variance changing over prescribed intervals. It was initially developed for the detection of change-points in daily climate data with annual periodic bias and variance changing on monthly intervals. It is quite versatile and can be of more general-purpose usage. The package includes the segmentation function as well as auxiliary functions for the screening, validation and visualization of the segmentation results. The theoretical basis of the method was published in [@Quarello2022]. 
 
## Table of Contents

- [Installation](#installation)
- [Examples](#examples)

## Installation

The development version of `PMLSeg` can be installed from GitHub with:

```{r, eval=FALSE}

# Install devtools and tidyr if you haven't already
install.packages("devtools")
install.packages("tidyr")

# Download and install gfpop from
https://cran.r-project.org/src/contrib/Archive/gfpop/gfpop_1.1.1.tar.gz
install.packages("gfpop_1.1.1.tar.gz")

# Install PMLseg from GitHub
devtools::install_github("khanhninhnguyen/PMLSeg")
```

## Functions

The package contains the following functions:

Segmentation(
  OneSeries,
  lmin = 1,
  Kmax = 30,
  selectionK = "BM_BJ",
  FunctPart = TRUE,
  selectionF = FALSE
)


PlotSeg(
  OneSeries,
  SegRes,
  FunctPart = TRUE,
  RemoveData = NULL,
  Metadata = NULL,
  Validated_CP_Meta = NULL,
  labelx = "Date",
  labely = "signal"
)

Cluster_screening(Tmu, 
    p_val = 0.05, 
    MaxDist = 80, 
    detail = NULL)

=> output:
 - UpdatedCP: The change-points remaining after screening.
 - RemoveData: A data frame containing the beginning and end positions (time index) of the segments to be deleted after filtering.
 - ChangeCP: "Yes" or "No", indicating whether the list of change-points changed after cluster screening.
 - detail: A data frame containing the results of the significance test of the mean difference before and after the cluster, including the following columns:
    mu_L, mu_R: Means of the segments before and after the cluster.
    se_L, se_R: Standard errors of the means of the segments before and after the cluster.
    np_L, np_R: Number of points before and after the cluster.
    tstat, pval, signif: t-values, p-values, and significance codes of the test.

UpdatedParametersForFixedCP(
  OneSeries,
  ResScreening,
  FunctPart = TRUE,
  selectionF = FALSE
)

Validation(
  OneSeries,        # data frame, n x 2, $signal, $dates in Date format
  Tmu,              # segmentation results
  MinDist = 62,     # maximum distance (in days) between detected CP in Tmu$ and known events in Metadata to be valid
  Metadata          # data frame with two columns: $date that is the known change in same format as date of OneSeries} and the $type that is the type the change
)

## Examples

Preview the examples given in the .md files.

The full .R codes are given in the folder Examples.

## References







