---
title: "Test PMLSeg"
author: "Ninh"
date: "2024-08-09"
output: beamer_presentation
font-family: 'Helvetica'
fontsize: 9pt
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Introduction

This document presents results from several tests of the `PMLSeg` package, including:

- Test of the `Segmentation` function with and without offsets, considering the following examples:

  + Ex1 : zero mean + IID noise
  
  + Ex2 : periodic mean + IID noise 
  
  + Ex3 : periodic mean + monthly variance 
  
- Tests of other functions such as:

  + `PlotSeg` to visualize segmentation results 
  
  + `Cluster_screening` to detect groups of close change-points (usually due to outliers) and determine whether to keep or remove the cluster. The cluster in this document is determined by a group of close change-points within 80 days (this threshold is changeable).
  
  + `Validation` to validate the detected change-points with the help of metadata.

## Generate Example Data 

Ex1 time series 

\footnotesize
```{r ex1, size = "small", echo = TRUE, fig.width = 6, fig.height = 3}
set.seed(1)
length_series = 1000
df1 = data.frame(date = seq.Date(from = as.Date("2010-01-01"),
                                to = as.Date("2010-01-01")+(length_series-1), 
                                by = "day"),  
                signal = rnorm(n = length_series, mean = 0, sd = 1))

plot(df1$date, df1$signal, type = "l")
head(df1, 3)
```

## Generate Example Data 

Ex2 time series : Add the functional with 4 Fourier series components, each with a coefficient of 0.5

\footnotesize
```{r ex2, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
library(dplyr)

T <- 365.25
df2 <- df1 %>%
  mutate(t = as.numeric(date - date[1])+1,
  f = rowSums(sapply(1:4, function(i) 0.5*cos(i*t*(2*pi)/T) + 0.5*sin(i*t*(2*pi)/T))),
  signal = signal + f)

head(df2, 3)
plot(df2$date, df2$f, type = "l")
```

## Generate Example Data 

Ex3 time series : Apply the same functional as Ex2, adjusted for monthly variance
\footnotesize
```{r ex3, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
std = c(1, 1.2, 1.4, 1.6, 1.8, 2, 2, 1.8, 1.6, 1.5, 1.4, 1.3)

df3 <- df1 %>% 
  mutate(sd = std[as.numeric(format(date, "%m"))],
         signal = signal * sd ) %>%
  mutate(signal = signal + df2$f)

head(df3, 3)
plot(df3$date, df3$signal, type = "l")
```

## Generate Example Data  

Harmonize Formats of 3 Dataframes for Testing

\footnotesize
```{r harmonize, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
df2 <- df2 %>% select(date, signal)
df3 <- df3 %>% select(date, signal)

names(df1)
names(df2)
names(df3)
```
## Preliminary Settings

Generate different offset series to add to the original series :
\footnotesize

```{r jump, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}

# Function to generate jump series
generate_jump_series <- function(jump_indices, jump_amp, length_series) {
  jump_series <- rep(0, length_series)
  jump_indices <- c(1, jump_indices, length_series + 1)
  
  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- jump_amp
  
  jump_series <- cumsum(changes)
  
  return(jump_series)
}
# No cluster
jump_ind1 <- c(200, 600)
jump_amp1 <- c(0, 1, 1)
# One cluster formed by the second and third change-points, which need to be 
# replaced by the middle point 
jump_ind2 <- c(200, 600, 640)
jump_amp2 <- c(0, 1, -2, 3)
# One cluster formed by the second and third change-points, which need to be removed 
jump_amp3 <- c(0, 1, -2, 2)
```

## Premilinary setting
\footnotesize

```{r jumpplot, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
# Generate jump series
jump_series1 <- generate_jump_series(jump_ind1, jump_amp1, length_series)
jump_series2 <- generate_jump_series(jump_ind2, jump_amp2, length_series)
jump_series3 <- generate_jump_series(jump_ind2, jump_amp3, length_series)
# Visualize jump series 
plot(jump_series1, type = "l", ylim = c(-2,3))
lines(jump_series2, col = "red")
lines(jump_series3, col = "blue")
```

## Premilinary setting
\footnotesize

```{r metadata, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
# Create an example metadata
meta = data.frame(date = c(df1$date[jump_ind2], df1$date[400]),  
                type = c("type1", "type2", "type3", "type3"))
meta
```
## Testing the `Segmentation` Function

When the series is homogeneous
\footnotesize
```{r testseg1a, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
library(PMLseg)
# Ex1 
seg1a = Segmentation(OneSeries = df1, FunctPart = FALSE)
str(seg1a)
seg1a$Tmu
```
\normalsize 
No change-points are detected, as shown in the `Tmu` dataframe, which lists all segments of the series. The mean is close to 0. Additionally, the results include not only the `Tmu`  dataframe but also other outputs such as the fitted functional part `FitF` (which is not fitted when `FunctPart` is set to FALSE), coefficients of functional elements `CoeefF`, monthly variance `MonthVar`, and the Sum of Squares of Residuals `SSR`. 

## Test the `Segmentation` Function

When series is homogeneous
\footnotesize
```{r testseg2a, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
# Ex2 
seg2a = Segmentation(OneSeries = df2, FunctPart = TRUE)
seg2a$Tmu
# Ex3 
seg3a = Segmentation(OneSeries = df3, FunctPart = TRUE)
seg3a$Tmu
```
\normalsize
No change-points are detected in these two examples either.

## Test the `Segmentation` Function

Adding Jump Series to the original Series and running the segmentation
\footnotesize
```{r testseg1b, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
# Ex1 
df1b <- df1 %>% mutate(signal = signal + jump_series1)
seg1b = Segmentation(OneSeries = df1b, FunctPart = FALSE)
seg1b$Tmu

# Ex2 
df2b <- df2 %>% mutate(signal = signal + jump_series1)
seg2b = Segmentation(OneSeries = df2b, FunctPart = TRUE)
seg2b$Tmu

# Ex3 
df3b <- df3 %>% mutate(signal = signal + jump_series1)
seg3b = Segmentation(OneSeries = df3b, FunctPart = TRUE)
seg3b$Tmu
```

## Test the `Segmentation` Function

Adding other jump series in the most complicated series (Ex3)

\footnotesize
```{r testseg3c, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}

df3c <- df3 %>% mutate(signal = signal + jump_series2)
seg3c = Segmentation(OneSeries = df3c, FunctPart = TRUE)
seg3c$Tmu

df3d <- df3 %>% mutate(signal = signal + jump_series3)
seg3d = Segmentation(OneSeries = df3d, FunctPart = TRUE)
seg3d$Tmu
```
## Test the `PlotSeg` Function

Visualize the segmentation result of example 1 with jump series 1
\footnotesize
```{r plotseg1b, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
PlotSeg(OneSeries = df1b, SegRes = seg1b, FunctPart = FALSE)
```

## Test the `PlotSeg` Function

Visualize the segmentation result of example 2 with jump series 1 
\footnotesize
```{r plotseg2b, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
PlotSeg(OneSeries = df2b, SegRes = seg2b, FunctPart = TRUE)
```

## Test the `PlotSeg` Function

Visualize the segmentation result of example 3 with with jump series 2
\footnotesize
```{r plotseg3c, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
PlotSeg(OneSeries = df3c, SegRes = seg3c, FunctPart = TRUE)
```

## Test the `PlotSeg` Function

Visualize the segmentation result of example 3 with with jump series 3
\footnotesize
```{r , echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
PlotSeg(OneSeries = df3d, SegRes = seg3d, FunctPart = TRUE)
```

## Test the `Cluster_screening` Function

Segmentation sometime detects the cluster like in example 3c, 3d. When it is too close and after cluster, the change in mean is not significant, we will remove such kind of cluster (`screening2`). If it is significant, we will replace a cluster by the middle point (`screening1`) and recommend to remove the data within the cluster. 

\footnotesize
```{r cluster, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
screening1 = Cluster_screening(Tmu = seg3c$Tmu,
                              MaxDist = 80)
str(screening1)

screening2 = Cluster_screening(Tmu = seg3d$Tmu,
                              MaxDist = 80)
str(screening2)
```

## Test the `UpdatedParametersForFixedCP` function

When the screening say it is needed to replace the changepoint, we need to reestimate the mean and other parameters in the `Tmu` dataframe. 

\footnotesize
```{r update, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
seg_upd = UpdatedParametersForFixedCP(OneSeries = df3c, 
                                      ResScreening = screening1)
str(seg_upd)
seg_upd$Tmu
```

## Test the `UpdatedParametersForFixedCP` function

Visualize the time series after the updated segmentation 

\footnotesize
```{r plotupdate, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 1.5}
PlotSeg(OneSeries = df3d, 
        SegRes = seg_upd,
        RemoveData = screening1$RemoveData)
```

Compare to the result before the cluster screening

```{r plotcompare, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 1.5}
PlotSeg(OneSeries = df3d, 
        SegRes = seg3d)
```

## Test the `Validation` function

Validate the detected change-points with respect to the metadata 

\footnotesize
```{r validation, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
valid = Validation(OneSeries = df3d, 
           Tmu = seg_upd$Tmu,
           MinDist = 62,
           Metadata = meta)
valid
```

## Test the `Validation` function

Visualize results with metadata

```{r plotmeta, echo = TRUE, message=FALSE, fig.width = 6, fig.height = 3}
PlotSeg(OneSeries = df3d, 
        SegRes = seg_upd,
        RemoveData = screening1$RemoveData,
        Metadata = meta, 
        Validated_CP_Meta = valid)
```
