---
output: github_document
---

<!-- Example1.md is generated from Example1.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 1: time series with 2 change-points and constant noise

### 1. Simulate a time series

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# Note 1: Note: by convention the date/time of a change-point is the date/time of the last point in a the segment

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n <- 1000                    # length of time series
cp_ind <- c(200, 600)       # position of CPs (index in time series)
segmt_mean <- c(-1, 1, 2)   # mean value of segments
noise_stdev <- 1             # noise std dev (identical for all months)
set.seed(1)                 # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df <- data.frame(date = mydate, signal = mysignal)

# plot signal and position of change-points (red dashed line)
plot(df$date, df$signal, type = "l", xlab = "Date", ylab = "signal")
abline(v = mydate[cp_ind], col = "red", lty = 2)
```

### 2. Segmentation

Run the segmentation with default parameters and no functional:

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
str(seg)
```

The `Tmu` dataframe contains, for each segment: the index of beginning and end, the esitmated mean and its standard erreor `se`, and the number of valid (non-NA) data points `np` in the signal:

```{r}
seg$Tmu
```

### 3. Visualization of the time series with segmentation results superposed

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE)
```

The plot shows the signal (grey line), the estimated means (red line), the estimated noise std (cyan line). 
The y-intercept of the noise is the lower black line. Dashed vertical lines show the estimated times of the change-points.

### 4. Validation of detected change-points with metadata

Metadata is represented by a data frame with 2 columns: `date`, `type`.

For the example, we create a fake metadata data frame with the true position of change-points:

```{r}
meta_ind = cp_ind               # index in time series of metadata information
meta_date <- df$date[meta_ind]  # corresponding date 
meta_type <- c("R", "RAD")      # type of information, e.g. R = receiver change, A = antenna change, D = radome change
metadata = data.frame(date = meta_date, type = meta_type)
metadata
```

Plot with metadata:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE, 
        Metadata = metadata) 
```

Validate estimated change-point positions wrt metadata:

```{r}
valid_max_dist = 10             # maximum distance wrt metadata for a CP to be validated
valid = Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```
Note: valid$Distance gives the distance between estimated CP and metadata

Plot with metadata and validation results:

```{r}
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = FALSE, Metadata = metadata, Validated_CP_Meta = valid)
```

Validated change-points are inducated by a filled triangle at the bottom line.

### 5. Further explore the sensitivity of segmentation results to signal and noise parameters

#### Impact of sample size on estimated parameters

  a) run again with `n = 4000` and observe that:
    - the position of change-points is the same as with `n = 1000`
    - the `seg$Tmu$mean` value the last segment (longer than with `n = 1000`) is more accurate and `seg$Tmu$se` is smaller
    - the `seg$MonthVar` values are closer to the true value (1)

  b) run again with `n = 500`

#### Impact of signal to noise ratio on estimated parameters

  c) run again with `noise_std = 0.01, 0.1` these are easy cases
  
  d) run again with `noise_std = 10` this is a difficult case





