---
output: github_document
---

<!-- Example6.md is generated from Example6.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 6: time series with periodic bias, monthly variance, and gaps

### 1. Simulate a time series 

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
    time_series <- rep(0, length_series)
    cp_indices <- c(1, cp_ind, length_series + 1)
    offsets <- c(0, diff(segmt_mean))

    changes <- rep(0, length_series)
    changes[cp_indices[-length(cp_indices)]] <- offsets
    changes[1] <- segmt_mean[1]
    time_series <- cumsum(changes)

    sd = noise_stdev[as.numeric(format(mydate, "%m"))]
    noise <- rnorm(n = length_series, mean = 0, sd = 1)
    time_series <- time_series + noise * sd

    return(time_series)
}

fourier_series <- function(mydate, coeff) {
    T <- 365.25                             # base period (unit days)
    t <- as.numeric(mydate)                 # time variable for periodic function
    t0 <- as.numeric(mydate[1])             # reference date of periodic function is first date of time series
    p <- length(coeff) / 2                  # order of Fourier Series
    f <- rowSums(sapply(1:p, function(i) coeff[2*i-1]*cos(i*(t-t0+1)*(2*pi)/T) + coeff[2*i]*sin(i*(t-t0+1)*(2*pi)/T)))

    return(f)
}

# specify the simulation parameters
date_begin <- as.Date("2010-03-01")             # date of first data point
n <- 1000                                        # length of time series
cp_ind <- c(200, 600, 990)                      # position of CPs (index in time series)
segmt_mean <- c(-1, 1, 2, -1)                    # mean value of segments
noise_stdev <- c(0.1, 0.3, 0.7, 1.2, 1.8, 2, 2, 1.8, 1.2, 0.7, 0.3, 0.1) # 12 values, one per month (Jan to Dec)
coeff <- c(1, 0, 0, 0)                          # Fourier Series coefficients (cos1, sin1, cos2, sin2...) up to order 4
set.seed(1)                                     # initialise random generator

# create a time series with jumps and noise
mydate <- seq.Date(from = date_begin, to = date_begin + n - 1, by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)

# add NA's in the signal
NA_ind = seq(from = 100, to = 150, by = 1)  # 1st gap
mysignal[NA_ind] <- NA
NA_ind = seq(from = 580, to = 630, by = 1)  # 2nd gap 
mysignal[NA_ind] <- NA

# add a periodic function (Fourier series)
f <- fourier_series(mydate, coeff)
mysignal <- mysignal + f

# create df with full signal
df <- data.frame(date = mydate, signal = mysignal)

# plot signal and position of change-points (red dashed line)
plot(df$date, df$signal, type = "l")
abline(v = mydate[cp_ind], col = "red", lty = 2)
```

### 2. Segmentation

#### a. Run the segmentation with default parameters:

```{r}
seg = Segmentation(OneSeries = df, FunctPart = TRUE)
seg$Tmu
seg$CoeffF
seg$MonthVar
seg$SSR
sum(seg$CoeffF^2)
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = TRUE)
```

#### b. run segmentation with selection of statistically significant Fourier coefficients

```{r}
seg_selectF = Segmentation(OneSeries = df, FunctPart = TRUE, selectionF = TRUE)
seg_selectF$Tmu
seg_selectF$CoeffF
seg_selectF$MonthVar
seg_selectF$SSR
sum(seg_selectF$CoeffF^2)
PlotSeg(OneSeries = df, SegRes = seg_selectF, FunctPart = TRUE)
```

### 3. Cluster screening 

It is safe to always run the cluster screening after the segmentation.

```{r}
# run the cluster screening
cluster_max_dist <- 80             # max distance between CPs in a cluster
screening <- Cluster_screening(Tmu = seg_selectF$Tmu, MaxDist = cluster_max_dist)
screening

# update the segmentation dataframe if CPs have changed
if (screening$ChangeCP == "Yes") {
    seg_updated <- UpdatedParametersForFixedCP(OneSeries = df, ResScreening = screening, FunctPart=TRUE)
} else {
    seg_updated <- seg_selectF
}
```

### 4. validate updated CP position wrt metadata

```{r}
true_cp_ind <- cp_ind
true_cp_df <- data.frame(date = df$date[true_cp_ind], type = rep("True", (length(true_cp_ind))))
valid_max_dist <- 10               # max distance between CP and metadata for the validation
valid <- Validation(OneSeries = df, Tmu = seg_updated$Tmu, MaxDist = valid_max_dist, Metadata = true_cp_df)
valid
```

### 5. Plot the time series
```{r}
PlotSeg(OneSeries = df, SegRes = seg_updated, FunctPart = TRUE, Metadata = true_cp_df, Validated_CP_Meta = valid, RemoveData = screening$RemoveData)
```

Remember that the data in the clusters are hidden in the Plot but are still in the time series dataframe. To remove them from the data see Example3.md.
