---
output: github_document
---

<!-- Example2.md is generated from Example2.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.width=6, fig.height=3
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 2: time series with missing data

### 1. Simulate time series

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n <- 1000                    # length of time series
cp_ind <- c(200, 600)        # position of change points (index in time series)
segmt_mean <- c(-1, 1, 2)    # mean value of segments
noise_stdev <- 1             # noise std dev (identical for all months)
set.seed(1)                  # initialise random generator

# simulate data
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)

# true CP date
CP_date <- mydate[cp_ind]                             # date of CP
CP_date

# define fake metadata from true CP dates
meta_date <- CP_date                                  # date of metadata event = date of CP
meta_type <- c("receiver_change", "antenna_change")   # type of metadata event
metadata = data.frame(date = meta_date, type = meta_type)

# add NA's in the signal
NA_ind <- seq(from = 100, to = 150, by = 1)  # 1st gap
mysignal[NA_ind] <- NA
NA_ind <- seq(from = 580, to = 630, by = 1)  # 2nd gap 
mysignal[NA_ind] <- NA

# create a time series df
df <- data.frame(date = mydate, signal = mysignal)

# plot signal and position of change-points (red dashed line)
plot(df$date, df$signal, type = "l", col = "gray", xlab = "date", ylab = "signal", main="Simulated time series")
abline(v = CP_date, col = "red", lty = 2)
```

Note that the 1st gap lies within the 1st segment, while the 2nd gap is overlapping the 2nd change-point.

### 2. Segmentation

Run the segmentation without the functional part:

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)
seg$Tmu
```

The first CP is detected at the right date but not the 2nd one because of the gap. However the missing data points do not count in `np`. 
Hence, the gap does not hamper the validation (see below).

### 3. Visualization of the time series with segmentation results and metadata superposed

Plot with metadata:

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE, 
        Metadata = metadata) 
```

### 4. Validation of detected change-points with metadata

```{r}
valid_max_dist = 10             # maximum distance wrt metadata for a CP to be validated
valid = Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```

Note that the distance between the 2nd CP (index=636) and the nearest metadata (index=600) excludes the NA values (up to 630): 636 - 631 = 5.

Hence both CPs are validated.

```{r}
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = FALSE, Metadata = metadata, Validated_CP_Meta = valid)
```


