---
output: github_document
---

<!-- Example3.md is generated from Example3.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.width=6, fig.height=3
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 3: time series with clusters of CPs

In real data, clusters of CPs can occur in a short period of time or can be due to noise spikes.
In many applications, we don't want to keep such clusters with short segments.
We have included a screening function which removes clusters or reduces them to single CPs, depending on the significance of the change in mean before/after each cluster. If the change is significant the cluster is replaced by a single CP, otherwise all the CPs in that cluster are removed.

### 1. Simulate a time series with 2 clusters

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]

  time_series <- cumsum(changes)
  noise <- rnorm(n = length_series, mean = 0, sd = noise_stdev)
  time_series <- time_series + noise

  return(time_series)
}

# specify the simulation parameters
n <- 1000                                       # length of time series
cp_ind <- c(10, 200, 210, 580, 590, 600, 990)   # 2 clusters of CPs + one short segment at the begining and one at the end
segmt_mean <- c(0, -1, 5, 1, -5, 5, 2, 0)       # mean of segments
noise_stdev <- 1                                # noise std dev (identical for all months)
set.seed(1)                                     # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
OneSeries <- data.frame(date = mydate, signal = mysignal)

CP_date <- mydate[cp_ind]                       # dates of CP

# plot signal and position of change-points
plot(OneSeries$date, OneSeries$signal, type = "l", col = "gray", xlab = "date", ylab = "signal", main="Simulated time series")
abline(v = CP_date, col = "red", lty = 2)
```

### 2. Segmentation

Run the segmentation without the functional part:

```{r}
SegRes = Segmentation(OneSeries = OneSeries, 
                   FunctPart = FALSE,
                   VarMonthly = FALSE)

print(SegRes$Tmu)
```

### 3. Visualization of the time series with segmentation results superposed

```{r}
PlotSeg(OneSeries = OneSeries, 
        SegRes = SegRes, 
        FunctPart = FALSE)
```

Note that the segmentation is able to detect all CPs, even those close to the beginning and end of the time series.

### 4. Cluster screening 

We want to remove the segments smaller than 80 days, either isolated or in clusters.

```{r}
screening <- Cluster_screening(Tmu = SegRes$Tmu, MaxDist = 80)
screening
```
The Cluster_screening function returns information to update the segmentation dataframe.

Now, update the segmentation parameters and the time serie

```{r}
SegResUpd <- UpdatedParametersForFixedCP(OneSeries = OneSeries, ResScreening = screening, FunctPart=FALSE, VarMonthly=FALSE)
print(SegResUpd)

OneSeriesUpd <- UpdateTimeSeries(OneSeries, screening$RemoveData)
```

Plot the updated time series with the updated segmentation information.

```{r}
PlotSeg(OneSeries = OneSeriesUpd, SegRes = SegResUpd, FunctPart = FALSE)
```


### 5. Validation of change-points after screening

```{r}
Metadata = data.frame(date = CP_date[c(2, 6)], type = c("1", "2"))
valid = Validation(OneSeries = OneSeriesUpd, Tmu = SegResUpd$Tmu, MaxDist = 10, Metadata = Metadata)
print(valid)

p <- PlotSeg(OneSeries = OneSeriesUpd, SegRes = SegResUpd, FunctPart = FALSE, Metadata = Metadata, Validated_CP_Meta = valid)
print(p)

```

