---
output: github_document
---

<!-- Example4.md is generated from Example4.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 4: time series with monthly varying variance

### 1. Simulate a time series with one different variance every calendar month 

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)
library(purrr)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]
  time_series <- cumsum(changes)

  noise <- rnorm(n = length_series, mean = 0, sd = 1)
  sd = noise_stdev[as.numeric(format(mydate, "%m"))]
  time_series <- time_series + noise * sd

  return(time_series)
}

# specify the simulation parameters
n <- 1000                            # length of time series
cp_ind <- c(200, 600)               # position of change points (index in time series)
segmt_mean <- c(-1, 1, 2)           # mean value of segments
noise_stdev <- c(0.1, 0.3, 0.7, 1.2, 1.8, 2, 2, 1.8, 1.2, 0.7, 0.3, 0.1) # 12 values, one per month (Jan to Dec)
set.seed(1)                         # initialise random generator

# create a data frame of time series with 2 columns: date, signal
mydate <- seq.Date(from = as.Date("2010-01-01"), to = as.Date("2010-01-01")+(n-1), by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)
df <- data.frame(date = mydate, signal = mysignal)

# plot signal and position of change-points (red dashed line)
plot(df$date, df$signal, type = "l",xlab ="Date",ylab="signal")
abline(v = mydate[cp_ind], col = "red", lty = 2)
```

### 2. Segmentation

Run the segmentation with default parameters:

```{r}
seg = Segmentation(OneSeries = df, 
                   FunctPart = FALSE)

seg$Tmu
```

### 3. Validate estimated change-point positions wrt metadata:

```{r}
meta_ind <- cp_ind               # index in time series of metadata information
meta_date <- df$date[meta_ind]  # corresponding date 
meta_type <- c("true", "true")      # type of information, e.g. R = receiver change, A = antenna change, D = radome change
metadata <- data.frame(date = meta_date, type = meta_type)
metadata

valid_max_dist <- 10             # maximum distance wrt metadata for a CP to be validated
valid <- Validation(OneSeries = df, 
           Tmu = seg$Tmu,
           MaxDist =  valid_max_dist,
           Metadata = metadata)
valid
```

### 4. Visualization of the time series with segmentation and validation results superposed

```{r}
PlotSeg(OneSeries = df, 
        SegRes = seg, 
        FunctPart = FALSE,
        Metadata = metadata, 
        Validated_CP_Meta = valid)
```

Note: the segmentation detects well the position of change points despite the steep variations in the noise 

### 5. Evaluate the error of the estimated noise variance
```{r}
error <- seg$MonthVar - noise_stdev ** 2
error / noise_stdev
```

The relative error on some months is quite large.

Increasing the length of the time series (i.e. including more years) would improve the accuracy of the noise variance estimation.
