---
output: github_document
---

<!-- Example5.md is generated from Example5.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Example 5: time series with periodic bias

### 1. Simulate a time series with periodic bias and constant noise variance

```{r}
rm(list=ls(all=TRUE))
library(PMLseg)
library(purrr)

# define simulation function
simulate_time_series <- function(cp_ind, segmt_mean, noise_stdev, length_series) {
  time_series <- rep(0, length_series)
  jump_indices <- c(1, cp_ind+1, length_series + 1)
  offsets <- c(0, diff(segmt_mean))

  changes <- rep(0, length_series)
  changes[jump_indices[-length(jump_indices)]] <- offsets
  changes[1] <- segmt_mean[1]
  time_series <- cumsum(changes)

  sd = noise_stdev[as.numeric(format(mydate, "%m"))]
  noise <- rnorm(n = length_series, mean = 0, sd = 1)
  time_series <- time_series + noise * sd

  return(time_series)
}

# specify the simulation parameters
date_begin <- as.Date("2010-03-01") # date of first data point
n <- 1000                            # length of time series
cp_ind <- c(200, 600)               # position of change points (index in time series)
segmt_mean <- c(-1, 1, 2)           # mean value of segments
noise_stdev <- rep(1, 12)         # noise variance (Jan to Dec)
coeff <- c(1, 0, 0, 0)              # Fourier Series coefficients (cos1, sin1, cos2, sin2...) up to order 4
set.seed(1)

# create a time series df with jumps and noise
mydate <- seq.Date(from = date_begin, to = date_begin + n - 1, by = "day")
mysignal <- simulate_time_series(cp_ind, segmt_mean, noise_stdev, n)

# create a periodic function (Fourier series)
T <- 365.25                             # reference period (unit days)
t <- as.numeric(mydate)                 # time variable for periodic function
t0 <- as.numeric(mydate[1])             # reference date of periodic function is first date of time series
p <- length(coeff) / 2                  # order of Fourier Series
f <- rowSums(sapply(1:p, function(i) coeff[2*i-1]*cos(i*(t-t0+1)*(2*pi)/T) + coeff[2*i]*sin(i*(t-t0+1)*(2*pi)/T)))

# add periodic function to the signal
mysignal <- mysignal + f
df <- data.frame(date = mydate, signal = mysignal)

# plot signal and position of change-points (red dashed line)
plot(df$date, df$signal, type = "l",xlab ="Date",ylab="signal")
abline(v = mydate[cp_ind], col = "red", lty = 2)
```

### 2. Segmentation

#### a. Run the segmentation without the functional part to see what happens...

```{r}
seg_nofunc <- Segmentation(OneSeries = df, FunctPart = FALSE)
seg_nofunc$Tmu
PlotSeg(OneSeries = df, SegRes = seg_nofunc, FunctPart = FALSE)
```

The segmentation captures the periodic signal.

#### b. Run the segmentation with the functional part.

```{r}
seg <- Segmentation(OneSeries = df, FunctPart = TRUE)
seg$Tmu
seg$CoeffF
seg$MonthVar
seg$SSR
sum(seg$CoeffF^2)
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = TRUE)
```

The segmentation now captures the CPs but the periodic signal is not well estimated by the functional part.

#### c. Run segmentation with functional and selection of statistically significant coefficients

```{r}
seg = Segmentation(OneSeries = df, FunctPart = TRUE, selectionF = TRUE)
seg$Tmu
seg$CoeffF
seg$MonthVar
seg$SSR
sum(seg$CoeffF^2)
PlotSeg(OneSeries = df, SegRes = seg, FunctPart = TRUE)
```

The functional part now correctly captures the periodic signal.